# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load the dataset and convert the date string to a date object.

```{r echo=TRUE}
act <- read.csv("activity.csv")
act$date <- as.Date(act$date, format = "%Y-%m-%d")
str(act)
```


## What is mean total number of steps taken per day?

Histogram of the number of steps taken each day:
```{r echo=TRUE}
hist(act$steps)
```

Mean and Median total number of steps taken per day:
```{r echo=TRUE}
meanSteps <- mean(act$steps[ ! is.na(act$steps)])
medianSteps <- median(act$steps[ ! is.na(act$steps)])
```
- Mean = `r meanSteps` 
- Median = `r medianSteps`

## What is the average daily activity pattern?

```{r echo=TRUE}
library(plyr)
meanInterval <- ddply(act[! is.na(act$steps),] ,~interval ,summarise,mean=mean(steps))
plot(meanInterval, type="l")
maxMeanInterval <- meanInterval[ max(meanInterval$mean), ]
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? `r maxMeanInterval`

## Imputing missing values

```{r echo=TRUE}
naRows <- is.na(act$steps)
naRowsCount <- sum(naRows)
```

Total number of missing values in the data set: `r naRowsCount`
Fill in missing data by simply inserting the average number of steps.

```{r echo=TRUE}
filledAct <- act
filledAct$steps[naRows] <- mean(filledAct$steps, na.rm = TRUE)
hist(filledAct$steps)
```

Comparison of mean and median between original data and filled in data
```{r echo=TRUE}
filledMean <- mean(filledAct$steps)
filledMedian <- median(filledAct$steps)
```

- Original Mean: `r meanSteps`
- Filled Mean: `r filledMean`

- Original Median: `r medianSteps`
- Filled Median: `r filledMedian`

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
library(ggplot2)
filledAct$weekend <- ifelse(weekdays(filledAct$date) %in% c("Saturdy","Sunday"), "Weekend", "Weekday")
groupedIntervals <- ddply( filledAct, weekend ~ interval, summarise, mean=mean(steps))
qplot(x=interval, y=mean, data=groupedIntervals, facets=weekend~., geom="line", ylab="Number of steps", xlab="Interval")
```



